Ages of Stars and Planets in the Kepler Field Over the First Four Billion Years

### 2024-08-28 papers/paper/ms.tex
update ms

% <250 words
\begin{abstract}
  %
  Recent analyses of FGK stars in open clusters have helped clarify
  the precision with which a star's rotation rate and lithium content
  can be used as empirical indicators for its age.
  %
  Here we apply this knowledge to stars observed by Kepler.
  %
  Rotation periods are drawn from previous work; lithium is measured
  from new and archival Keck/HIRES spectra.
  %
  We report rotation-based ages for \nuniqstarsantosrotgyroappl\ stars
  (harboring \nplwgyroagenograzing\ known planets) for which our method is
  applicable.
  %
  We find that our rotational ages accurately recover the ages of
  stars in open clusters spanning 0.04-2.5\,Gyr; they also agree with
  $\gtrsim$90\% of the independent lithium ages.
  %
  The resulting yield includes \nplyounggyrotwosigma\ planets younger
  than 1\,Gyr at 2$\sigma$, and \nplyounggyro\ with median ages below
  1\,Gyr.
  %
  This is about half the number expected under the classic assumption
  of a uniform star formation history.  We find that the scarcity of
  sub-gigayear systems can be attributed to the star formation rate in
  the Kepler field \replaced{dropping}{smoothly decreasing} by a factor of \replaced{\ratiosfr$\pm$\uncratiosfr}{a few}
  over the past 3\,Gyr.  We observe this trend both for known planet
  hosts and for the parent stellar sample.
  %
  This ``demographic cliff'' in the Galaxy's star formation history
  has been previously reported, and its confirmation helps clarify the
  age distribution of the known transiting exoplanets.
  %
\end{abstract}

>>>>

% <250 words
\begin{abstract}
  %
  Recent analyses of FGK stars in open clusters have helped clarify
  the precision with which a star's rotation rate and lithium content
  can be used as empirical indicators for its age.
  %
  Here we apply this knowledge to stars observed by Kepler.
  %
  Rotation periods are drawn from previous work; lithium is measured
  from new and archival Keck/HIRES spectra.
  %
  We report rotation-based ages for \nuniqstarsantosrotgyroappl\ stars
  (harboring \nplwgyroagenograzing\ known planets) for which our method is
  applicable.
  %
  We find that our rotational ages accurately recover the ages of
  stars in open clusters spanning 0.04-2.5\,Gyr; they also agree with
  $\gtrsim$90\% of the independent lithium ages.
  %
  The resulting yield includes \nplyounggyrotwosigma\ planets younger
  than 1\,Gyr at 2$\sigma$, and \nplyounggyro\ with median ages below
  1\,Gyr.
  %
  This is about half the number expected under the classic assumption
  of a uniform star formation history.  We find that the scarcity of
  sub-gigayear systems can be attributed to the star formation rate in
  the Kepler field \replaced{dropping}{smoothly decreasing} by a factor of \replaced{\ratiosfr$\pm$\uncratiosfr}{3-5}
  over the past 3\,Gyr.  We observe this trend both for known planet
  hosts and for the parent stellar sample.
  %
  This ``demographic cliff'' in the Galaxy's star formation history
  has been previously reported, and its confirmation helps clarify the
  age distribution of the known transiting exoplanets.
  %
\end{abstract}
